---
title: "DS311 - R Lab Assignment"
author: "Ann Yang"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
library(magrittr)
install.packages("dplyr")
library(dplyr)
library(ggplot2)

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
dim(mtcars)

# Answer:
print("There are total of 11 variables and 32 observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)
str(mtcars)

# Answer:
print("There are 0 discrete variables and 11 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m <- mean(mtcars$mpg)
v <- var(mtcars$mpg)
s <- sd(mtcars$mpg)


print(paste("The average of Mile Per Gallon from this data set is ", m , " with variance ", v , " and standard deviation", s , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
library(dplyr)
mtcars %>% 
  group_by(cyl) %>%
  summarize(Avgmpg = mean(mpg))

mtcars %>% 
  group_by(gear) %>%
  summarize(stdvmpg = sd(mpg))

```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
crosstab <- table(mtcars$cyl, mtcars$gear)
 crosstab
print("The most common car type in this data set is car with 3 cylinders and 8 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
library(ggplot2)
ggplot(data = PlantGrowth, aes(x = group, y = weight, fill = group)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Weight for Different Groups",
       x = "Group",
       y = "Weight")

```

Result:

=> Report a paragraph to summarize your findings from the plot!

treatment 1 has a negative effect on weight of plants while treatment 2 has a positive effect on the weight of plants. The weight of the control group was 50 while treatment 1 was around 45 and treatment 2 was around 55

b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}

ggplot(data = mtcars) + geom_histogram(aes(x = mpg))


print("Most of the cars in this data set are in the class of 16 mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
ggplot(USArrests, aes(x = Murder, y = Assault)) +
  geom_point() +
  labs(title = "Scatter Plot of Murder and Assault",
       x = "Murder",
       y = "Assault")

pairs(USArrests, main = "Pairs Plot for USArrests Data Set")


```

Result:

=> Report a paragraph to summarize your findings from the plot!
There is positive correlation between murder and assault. As the number of murder increases so does the number of assaults and vice-versa.


***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
housingData %>% 
  group_by(Boro) %>%
  summarize(AvgYear=mean(Year.Built), AvgMarketValuePerSquareFeet=mean(Market.Value.per.SqFt))

housingData %>% 
  group_by(Year.Built) %>%
  summarize(CountOfHouses= n(), AvgMarketValuePerSquareFeet=mean(Market.Value.per.SqFt))
```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
ggplot(housingData, aes(x = Year.Built, y = Market.Value.per.SqFt)) +
    geom_point(aes(color=Boro)) +
  labs(title = "Scatter Plot of market value per square foot by year built",
       x = "Year Built",
       y = "Market Value per Square Foot")

ggplot(housingData, aes(x = Boro, y = Year.Built)) +
  geom_boxplot() +
  labs(title = "Scatter Plot of Boro by Year Built",
       x = "Boro",
       y = "Year Built")
```

c. Write a summary about your findings from this exercise.
 
=> Enter your answer here!

In graph one we can find that
  - Manhattan 
    - has the highest market value pe square foot through the years
    - Most homes in Manhattan were built in 1960
  - Around 1950 there were fewer homes being built
  - Most of the homes built in Queens was after 1925 and mostly around the 2000s
  - Bronx
    - Market value pesquare foot is the lowest
    - Most Homes were built in 2000
  - Very few homes were built in Brooklyn around 1950


